---
title: "Analysis"
output: html_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(ggplot2)
library(rstatix)
library(knitr)
library(gganimate)
library(png)
library(gifski)
library(ARTool)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)
```

```{r data, echo = FALSE, message = FALSE}

#load dataset
nyc_df = 
  read_csv(file = "./data/New_York_City_Leading_Causes_of_Death.csv") %>%
  janitor::clean_names() 

write_csv(nyc_df, "./data/New_York_City_Leading_Causes_of_Death.csv")

# remove parentheses from leading cause column

nyc_df$leading_cause = 
  gsub("\\s*\\([^\\)]+\\)","",as.character(nyc_df$leading_cause))

```

## **Leading Causes of Death**

### What are the leading causes of death in the US?

```{r us plot 1, echo = FALSE, error = TRUE, message = FALSE}
nyc_df_1 = read_csv(file = "./data/nyc_clean.csv")
US_df_1 = read_csv(file = "./data/Leading_Causes_of_Death_States_Only.csv")

# to create ranking
US_df_2 = US_df_1 %>%
          subset(US_df_1$x113_cause_name != "All Causes") %>%
          group_by(x113_cause_name, year) %>% 
          summarise(total_death = sum(deaths)) %>% 
          arrange(desc(total_death)) %>% 
          group_by(year) %>% 
          mutate(order = min_rank(total_death) * 1.0) %>% 
          ungroup()

options(scipen=999)

#animated plot for US! 
anim_us = US_df_2 %>%
  ggplot(aes(order, group = x113_cause_name)) +
  geom_tile(aes(y = total_death / 2, height = total_death, width = 0.9, fill = x113_cause_name)) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
  scale_y_continuous(name = "Total Death Count") +
  transition_states(year, transition_length = 4, state_length = 3) +
  geom_text(aes(y = 0, label = x113_cause_name), hjust = 1) +
  coord_flip(clip = "off") +
  ggtitle("Top Causes of Death in US by Year: {closest_state}")+
  theme(legend.position = "none", 
        plot.margin = margin(1,1,1,7.5, 'cm'), 
        title = element_text(size = 9))

animate(anim_us, nframes = 100)

#original static plot we made for US 
#US_df_1_plot = 
  #US_df_1 %>%
  #filter(x113_cause_name != "All Causes") %>% 
  #group_by(x113_cause_name) %>%
  #summarize(total_death = sum(deaths)) %>%
  #arrange(desc(total_death)) %>%
  #mutate(x113_cause_name = fct_reorder(x113_cause_name, total_death)) %>% 
  #ggplot(aes(x = x113_cause_name, y = total_death, fill = x113_cause_name)) +
  #geom_col() +
  #labs(title = "Top 10 Leading Causes of Death in the U.S.", 
    #x = "Cause of Death", 
    #y = "Number of Deaths") +
  #theme(legend.position = "none") +
  #coord_flip() 

```

* It appears that the top 2 leading causes of death in the U.S. (#1 Diseases of heart and #2 Malignant neoplasms) remained the same from 1999 to 2017 with some variations in the rest of the list.

### What are the leading causes of death in NYC?

```{r plot 1, echo = FALSE, error = TRUE, message = FALSE}
nyc_df_1 = read_csv(file = "./data/nyc_clean.csv")

# to create ranking
nyc_df_2 = nyc_df_1 %>%
          subset(nyc_df_1$leading_cause != "All Other Causes") %>%
          group_by(leading_cause, year) %>% 
          summarise(total_death = sum(deaths)) %>% 
          arrange(desc(total_death)) %>% 
          subset(total_death > 200) %>%
          group_by(year) %>% 
          mutate(order = min_rank(total_death) * 1.0) %>% 
          ungroup()

options(scipen=999)

#animated plot for nyc! 
anim_nyc = nyc_df_2 %>%
  ggplot(aes(order, group = leading_cause)) +
  geom_tile(aes(y = total_death / 2, height = total_death, width = 0.9, fill = leading_cause)) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
  scale_y_continuous(name = "Total Death Count") +
  transition_states(year, transition_length = 4, state_length = 3) +
  geom_text(aes(y = 0, label = leading_cause), hjust = 1) +
  coord_flip(clip = "off") +
  ggtitle("Top Causes of Death in NYC by Year: {closest_state}")+
  theme(legend.position = "none", 
        plot.margin = margin(1,1,1,7.5, 'cm'), 
        title = element_text(size = 9))

animate(anim_nyc, nframes = 100)

#original static plot we made for NYC
#nyc_df_1 %>%
  #group_by(leading_cause) %>%
  #summarize(total_death = sum(deaths, na.rm = TRUE)) %>%
  #arrange(desc(total_death)) %>%
  #head(10) %>%
  #mutate(leading_cause = fct_reorder(leading_cause, total_death)) %>% 
  #ggplot(aes(x = leading_cause, y = total_death, fill = leading_cause)) +
  #geom_col() +
  #labs(title = "Top 10 Leading Causes of Death in NYC", 
    #x = "Cause of Death", 
    #y = "Number of Deaths") +
  #theme(legend.position = "none") +
  #coord_flip() 
  
```

* It appears that the top 2 leading causes of death in New York City (#1 Diseases of heart and #2 Malignant neoplasms) remained the same from 2007 to 2014 with a lot of variations in the rest of the list. 
* This high variability in the total death count for other causes of death could be due to unique epidemiologic factors in New York City or changes in the causes of death that were measured throughout the years in New York City. 

```{r, echo = FALSE, message = FALSE}
# Death rate change for US

us_df = read_csv(file = "./data/Leading_Causes_of_Death_US_Only.csv", col_types = "ffffid") %>%
  filter(cause_name != "All causes") 

us_deaths = us_df %>%
  group_by(x113_cause_name) %>%
  summarize(Counts = sum(deaths)) %>% 
  arrange(Counts) %>% 
  tail(3)
  
# Top 3 leading causes of death for US from 1999 to 2017 are: 
# "Diseases of heart"
# "Malignant neoplasms"
# "Cerebrovascular diseases"
```

## **Changes in the Leading Causes of Death**

### How has the death count of top leading causes changed over the years in the US?
```{r us plot 2, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE}
US_df_1_plot_1 = 
  US_df_1 %>% 
  filter(x113_cause_name == c("Diseases of heart", "Malignant neoplasms", "Cerebrovascular diseases")) %>% 
  group_by(year, x113_cause_name) %>%  
  summarize(total_death = sum(deaths)) %>%
  arrange(desc(total_death)) %>%
  mutate(x113_cause_name = fct_reorder(x113_cause_name, total_death)) %>% 
  ggplot(aes(x = year, y = total_death, color = x113_cause_name)) + 
  geom_line(size = 1) + 
  labs(title = 'Top 3 Leading Causes of Death in US (2000-2015)', x = 'Year', y = 'Death Count') + 
  transition_reveal(year)

US_df_1_plot_1
```

* It appears that from 1999 to 2017, deaths due to malignant neoplasms in the U.S. showed a steadily increasing trend, while deaths due to diseases of heart showed a declining trend initially but started increasing again in recent years. Deaths due to cerebrovascular diseases remained relatively stable throughout the years.

### How has the death count of top 2 leading causes changed over the years in NYC?

```{r plot 2, echo = FALSE, error = TRUE, message = FALSE}
nyc_change = 
  nyc_df_1 %>%
  filter(leading_cause %in% c("Malignant Neoplasms", "Diseases of Heart")) %>%
  mutate(deaths = as.numeric(deaths)) %>%
  group_by(year, leading_cause) %>%  
  summarize(total_death = sum(deaths)) %>%
  ggplot(aes(x= year, y = total_death, color = leading_cause)) + 
  geom_line(size=1) + 
  labs(title='Top 2 Leading Causes of Death in NYC (2007-2014)', x='Year', y='Death count')+ 
  theme(legend.text= element_text(size =8))+ 
  transition_reveal(year)

nyc_change
```

* It appears that from 2007 to 2014, deaths due to malignant neoplasms in NYC also showed an increasing trend but the change is not as drastic as in the nationwide plot, while deaths due to diseases of heart showed a declining trend, similar to the U.S. plot.  
* We originally wanted to show changes in top 3 leading causes for NYC as well similar to the US time series plot, but a lot of the data was fragmented (e.g. not available for all the years) for causes other than "All Other Causes" and we didn't think it was meaningful to include "All Other Causes" as one of the top causes for this plot, so we decided to just include "Malignant Neoplasms" and "Diseases of Heart". 

## **Racial and Sex Differences in NYC Mortality**

### How have the top 2 leading causes of death changed for each race in NYC?


```{r plot 3, echo = FALSE, error = TRUE, message = FALSE}

nyc_df_1 %>%
  filter(leading_cause %in% c("Malignant Neoplasms", "Diseases of Heart")) %>%
  mutate(deaths = as.numeric(deaths)) %>%
  group_by(year, leading_cause, race_ethnicity) %>%  
  summarize(total_death = sum(deaths)) %>%
  ggplot(aes(x= year, y = total_death, color = leading_cause)) + 
  geom_line(size=1) + 
  labs(title='Top 2 leading causes of death by ethnicity in NYC (2007-2014)', x='Year', y='Death count') +
  facet_wrap(~race_ethnicity, ncol=3, scales = 'free_y')

```

* It appears that the general trends in the top 2 leading causes in NYC varied by race/ethnicity. One noticeable contrast that stands out is that deaths due to malignant neoplams seem to be on a steeper increasing trend for Asians/Pacific Islanders and Other race/ethnicity in NYC unlike for rest of the ethnicity groups. 
* Another noticeable difference is that deaths due to diseases of heart seem to be on an increasing trend for Other race/ethnicity in NYC unlike for rest of the ethnicity groups.

### Are there differences in the leading causes of death between men and women in NYC?

```{r plot 4, echo = FALSE, error = TRUE, message = FALSE}

nyc_df_1 %>%
  filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms", "Diabetes Mellitus", "Chronic Lower Respiratory Diseases")) %>% 
  group_by(leading_cause, sex) %>% 
  summarize(total_death = sum(deaths, na.rm = TRUE)) %>%
  ggplot(aes(x = sex, y = total_death, fill = sex)) +
  geom_col() +
  labs(title='Gender differences in leading causes of death in NYC', x='Sex', y= 'Death count') +
  facet_wrap(~leading_cause, scales = 'free_y')

```

* While there seem to be some visible gender differences in these leading causes of deaths in NYC, it is unclear whether these differences are statistically significant solely based on these plots. We further investigated these differences in our statistical analyses section below. 

## **Statistical Analyses**

### Is there a statistically significant difference in death rates between men and women in New York City?

__Summary statistics__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  group_by(sex) %>%
  get_summary_stats(death_rate, type = "common") %>%
  kable()
```

__Visualization__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  ggplot(aes(x = sex, y = death_rate)) +
  geom_boxplot() +
  labs(x='Sex', y= 'Death rate')
```

__Assumptions__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  group_by(sex) %>%
  shapiro_test(death_rate) %>%
  kable()

```

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_male = nyc_df_1 %>%
  filter(sex == "M") %>%
  select(death_rate) 

  qqnorm(nyc_male$death_rate) 
  qqline(nyc_male$death_rate) 
```

From the boxplot visualization and the p-values (< 0.05) associated with the The Shapiro–Wilk test, there is evidence that the data from both the sex groups is not normally distributed. This is also evident from the qqplot above.     

Since the normality assumption has been violated, the Wilcoxon rank sum test (non-parametric) will be performed. 

__Test__

```{r stat_test1, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}

nyc_df_1 %>% 
  wilcox_test(death_rate ~ sex) %>%
  add_significance() %>%
  kable()
```

p-value = 0.00125 < 0.05

__Report__

The respective median death rates for male and female groups are as follows: 20 and 17.45. Based on the Wilcoxon rank sum test result, we have evidence to state that there is a statistically significant difference in death rates between men and women in New York City (p = 0.00125). 
        
            
### Is there a statistically significant difference in death rates across different ethnicities in New York City?    
       
__Summary statistics__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  mutate(race_ethnicity = as.factor(race_ethnicity)) %>%
  group_by(race_ethnicity) %>%
  summarize(mean = mean(death_rate), 
            sd = sd(death_rate), 
            median = median(death_rate), 
            IQR = IQR(death_rate, na.rm = T)) %>%
  kable()
```

Since the "Other" and "Unknown" categories of race don't have any associated death rates, these categories have been filtered out. 

__Visualization__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  filter(!race_ethnicity == "Other" & !race_ethnicity == "Unknown") %>%
  ggplot(aes(x = race_ethnicity, y = death_rate)) +
  geom_boxplot() +
  labs(x='Race', y= 'Death rate')
```

__One-way ANOVA test__

```{r anova, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE}
anova_df = nyc_df_1 %>%
           filter(!race_ethnicity == "Other" & !race_ethnicity == "Unknown")

anova_test(death_rate ~ race_ethnicity, data = anova_df) %>% kable()
```
        
          
p-value = 1.83e-14 < 0.05

At 0.05 significance level, we reject the null hypothesis and conclude that there is a significant difference in death rates across different ethnicities, in New York City.

__Multiple Comparisons__

A Tukey pairwise-comparisons test is performed to account for multiple comparisons
```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
mod.aov = aov(death_rate ~ race_ethnicity, data = anova_df)

TukeyHSD(mod.aov) 
```

From the above table, it can be inferred that there was a significant difference in death rates for the following pairs: Black and Asian/PI, White-Asian/PI, Hispanic-Black, White-Hispanic. 

__Assumptions__

Homogeneity of variances   
    
```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
levene_test(death_rate ~ race_ethnicity, data = anova_df) %>% kable()
```

Levene’s test was performed to check for the homogeneity of variances assumption. Based on the p-value, it can be inferred that there is no homogeneity of variances in different ethnicities.  

Normality

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
mod_residuals = residuals(object = mod.aov)

shapiro.test(x = mod_residuals)
```

Since the p-value < 0.00000000000000022, it can be assumed that the residuals are not normally distributed. Therefore, a Kruskal-Wallis rank sum test will be performed because the assumptions of one-way ANOVA have been violated. 

__Kruskal-Wallis rank sum test__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
kruskal.test(death_rate ~ race_ethnicity, data = anova_df)
```

__Report__

Based on the Kruskal-Wallis rank sum test result, we have evidence to state that there is a statistically significant difference in death rates across different ethnicities in New York City (p-value < 0.00000000000000022). 

### Are there significant differences in the leading causes of death between men and women in NYC?  

__Summary statistics__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms", "Diabetes   Mellitus", "Chronic Lower Respiratory Diseases")) %>% 
  group_by(leading_cause, sex) %>%
  summarize(count = n(),
            mean = mean(deaths), 
            sd = sd(deaths), 
            median = median(deaths)) %>%
  kable()
```

A slight inequality can be observed in the sample size per group. 

__Visualization__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms", "Diabetes   Mellitus", "Chronic Lower Respiratory Diseases")) %>% 
  group_by(leading_cause, sex) %>%
  ggplot(aes(x = leading_cause, y = deaths, color = sex)) +
  geom_boxplot() +
  labs(x='Leading Causes of Death', y= 'Death count')
```

__Two-way ANOVA test__

```{r , echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
two_anova_df = nyc_df_1 %>%
           filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms",  "Diabetes   Mellitus", "Chronic Lower Respiratory Diseases")) %>%
           mutate(
           sex = as.factor(sex),
           leading_cause = as.factor(leading_cause))

mod_two_anova = aov(deaths ~ leading_cause + sex, data = two_anova_df)

Anova(mod_two_anova, type = "III")
```
  
Two-way ANOVA test with interaction effect

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
mod_two_anova_2 = aov(deaths ~ leading_cause * sex, data = two_anova_df)

Anova(mod_two_anova_2, type = "III")
```

At 0.05 significance level, the following inferences can be made based on the p-values given above:

* The leading causes of deaths (Diseases of Heart, Malignant Neoplasms, Diabetes   Mellitus and Chronic Lower Respiratory Diseases) are associated with significant death counts in the New York City (p = 0.000001394). 
* The sex variable and the interaction term (sex and leading causes of death variable) don't have any significant association with death counts in New York City

__Assumptions__

Homogeneity of variances   
    
```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
levene_test(deaths ~ leading_cause*sex, data = two_anova_df) %>% kable()
```

Levene’s test was performed to check for the homogeneity of variances assumption. Based on the p-value, it can be inferred that there is no homogeneity of variances in the observed groups.  

Normality

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
mod_residuals_2 = residuals(object = mod_two_anova_2)

shapiro.test(x = mod_residuals_2)
```

Since the p-value = 0.000000000000001976, it can be assumed that the residuals are not normally distributed. Therefore, a Aligned Rank Transformed ANOVA will be performed because the assumptions of two-way ANOVA have been violated. 

__Aligned Rank Transformed ANOVA for two-way__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
art_df = two_anova_df %>% 
  drop_na(deaths) %>% 
  select(leading_cause, sex, deaths)

mod_art = art(deaths ~ leading_cause*sex, data = art_df)
anova(mod_art)
```

__Report__

Based on the Aligned Rank Transformed ANOVA for two-way procedure result, there is evidence to state that the leading causes of deaths (Diseases of Heart, Malignant Neoplasms, Diabetes   Mellitus and Chronic Lower Respiratory Diseases) is the only factor associated with significant death counts in the New York City (p = 0.00000000018816). 

### Are there significant differences in the top 2 leading causes of death across different ethnicities in NYC?  

__Summary statistics__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms")) %>% 
  group_by(leading_cause, race_ethnicity) %>%
  summarize(count = n(),
            mean = mean(deaths), 
            sd = sd(deaths), 
            median = median(deaths)) %>%
  kable()
```

__Visualization__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
nyc_df_1 %>%
  filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms")) %>% 
  group_by(leading_cause, race_ethnicity) %>%
  ggplot(aes(x = race_ethnicity, y = deaths, color = leading_cause)) +
  geom_boxplot() +
  labs(x='Ethnicities', y= 'Death count', color = "Leading causes")
```

__Two-way ANOVA test__

```{r , echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
two_anova_df_2 = nyc_df_1 %>%
           filter(leading_cause %in% c("Diseases of Heart","Malignant Neoplasms")) %>%
           mutate(
           race_ethnicity = as.factor(race_ethnicity),
           leading_cause = as.factor(leading_cause))

mod_two_anova_3 = aov(deaths ~ race_ethnicity + leading_cause, data = two_anova_df_2)

summary(mod_two_anova_3)
```
  
Two-way ANOVA test with interaction effect

```{r , echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
mod_two_anova_4 = aov(deaths ~ race_ethnicity*leading_cause, data = two_anova_df_2)

summary(mod_two_anova_4)
```

At 0.05 significance level, the following inferences can be made based on the p-values given above:

* The leading causes of deaths (Diseases of Heart and Malignant Neoplasms) are associated with significant death counts in the New York City (p <0.0000000000000002).
* The different categories of ethnicities are associated with significant death counts in the New York City (p <0.0000000000000002).
* The interaction term (ethnicities and leading causes of death variable) is associated with significant death counts in the New York City (p <0.0000000000000002).  

__Assumptions__

Homogeneity of variances   
    
```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
levene_test(deaths ~ race_ethnicity*leading_cause, data = two_anova_df_2) %>% kable()
```

Levene’s test was performed to check for the homogeneity of variances assumption. Based on the p-value, it can be inferred that there is no homogeneity of variances in the observed groups.  

Normality

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
mod_residuals_4 = residuals(object = mod_two_anova_4)

shapiro.test(x = mod_residuals_4)
```

Since the p-value < 0.00000000000000022, it can be assumed that the residuals are not normally distributed. Therefore, a Aligned Rank Transformed ANOVA will be performed because the assumptions of two-way ANOVA have been violated. 

__Aligned Rank Transformed ANOVA for two-way__

```{r, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE, comment = ""}
art_df_2 = two_anova_df_2 %>% 
  drop_na(deaths) %>% 
  select(leading_cause, race_ethnicity, deaths)

mod_art_2 = art(deaths ~ race_ethnicity*leading_cause, data = art_df_2)
anova(mod_art_2)
```

__Report__

Based on the Aligned Rank Transformed ANOVA for two-way procedure result, there is evidence to state that all the factor (race_ethnicity, leading_cause and interaction term) are associated with significant death counts in the New York City (p < 0.000000000000000222). Therefore, it is safe to state that there are significant differences in the top 2 leading causes of death across different ethnicities in NYC.  

```{r, echo = FALSE, message = FALSE, error = TRUE}


# find top 10 leading causes for each state each year from 1999 - 2017
states_df = read_csv(file = "./data/Leading_Causes_of_Death_States_Only.csv", col_types = "ffffid") %>%
  filter(cause_name != "All Causes") %>%
  arrange(year, state, age_adjusted_death_rate) %>%
  slice(which(row_number() %% 10 == 0))

# mode function
mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

states_summary =   
  states_df %>% group_by(year) %>% 
  summarize(Mode = mode(cause_name))

```


```{r us plot 3, echo = FALSE, error = TRUE, message = FALSE, warning = FALSE}
#US_df_1_plot_2 = 
#  US_df_1 %>% 
#  filter(x113_cause_name == c("Diseases of heart", "Malignant neoplasms", "Cerebrovascular diseases")) %>% 
#  group_by(x113_cause_name, state) %>%  
#  summarize(total_death = sum(deaths)) %>%
#  arrange(desc(total_death)) %>%
#  ggplot(aes(x = state, y = total_death, color = x113_cause_name)) + 
#  geom_line(size = 1) + 
#  facet_wrap(~x113_cause_name, ncol=3, scales = 'free_y') +
#  labs(title = 'Top 3 Leading Causes of Death by States', x = 'State', y = 'Death Count') 

```

