---
title: "Data"
output: html_document
---

For this project, two datasets were used: 


### 1) [New York City Leading Causes of Death](https://data.cityofnewyork.us/Health/New-York-City-Leading-Causes-of-Death/jb7j-dtam) <br> 

#### **Data Description** 
New York City Leading Causes of Death contains the leading causes of death by sex and ethnicity in New York City since 2007. Cause of death is derived from the NYC death certificate which is issued for every death that occurs in New York City.

```{r, include = FALSE}
library(tidyverse)
library(readxl)

nyc_df = read_csv(file = "./data/nyc_clean.csv")
```

##### The file contains a single dataframe with `r nrow(nyc_df)` rows of data on `r ncol(nyc_df)` variables: 

*   `r names(nyc_df[1])`
*   `r names(nyc_df[2])`
*   `r names(nyc_df[3])`
*   `r names(nyc_df[4])`
*   `r names(nyc_df[5])`
*   `r names(nyc_df[6])`
*   `r names(nyc_df[7])`. 

### 2) [Leading causes of deaths in US at the state-level](https://data.cdc.gov/NCHS/NCHS-Leading-Causes-of-Death-United-States/bi63-dtpu)

#### **Data Description**
This dataset presents the age-adjusted death rates for the 10 leading causes of death in the United States beginning in 1999. Data are based on information from all resident death certificates filed in the 50 states and the District of Columbia using demographic and medical characteristics. 

```{r, include = FALSE}
US_df = read_csv(file = "./data/Leading_Causes_of_Death_States_Only.csv")
```

##### The file contains a single dataframe with `r nrow(US_df)` rows of data on `r ncol(US_df)` variables: 

*   `r names(US_df[1])`
*   `r names(US_df[2])`
*   `r names(US_df[3])`
*   `r names(US_df[4])`
*   `r names(US_df[5])`
*   `r names(US_df[6])`
